---
title: "Aides Drug"
author: "Rabi Elkhoury"
date: "7/12/22"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
females <- filter(triUmp, gender == "F")
```

```{r eval=FALSE}
females <- females %>%
mutate(across(where(is.character), as.factor))
```


# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Does the administering of a new drug (triUmp) meant to stop AIDS infection to females actually prove effective?

$H0$ There is no difference between taking the drug and taking the placebo for females getting aids.
$HA$ There is a difference between taking the drug and taking the placebo for females getting aids.

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

To filter the data down to only females, we'll need to use a command available via dplyr to separate the classes.  From there, we can run graphical, numerical, and inferential tests to determine whether or not there is a tangible difference in AIDS numbers among those who were and were not treated with the AIDS drug.


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```


When the barplot(s) are examined, it's quite clear that most of the female population did not get aids, whether or not they had taken the drug or placebo.  However, for the females who did get aids, it seems that a majority of them had taken the drug rather than the placebo-based alternative.  This might imply that the drug is largely ineffective in protecting against the syndrome, but more testing is required to corroborate this.  


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```

Numerically, it's clear that the majority of people of took the drug did not get AIDS.  Even among those who took the placebo, only 2.42% of the female population got AIDS.  However, examining the data of those infected with AIDS supports the results generated in the earlier barplot.  For the females who were infected with AIDS, 58.82% had taken the drug meant to lessen the likelihood of infection.  This suggests that the drug itself is not effective in its aim.  Further inferential testing is necesitated to confirm this.



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Analyzing the Chi-Square test generated above, we find the p-value to be .0001; because this is lesser than .05, we reject the null hypothesis in favor of the alternate hypothesis.  This provides the evidence to suggest that there IS a difference between the females persons who did and did not receive the medication.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

There does seem to be a difference between among the female population who takes the drug, but it subverts reason, particularly because adminstering the drug seems to CAUSE a higher incidence of AIDS among females.  In advising a female patient as to whether or not they should take the medicine, counseling against it would be wisest, given the percentage/counts-based numerical and graphical data corroborating one another, along with the decisive Chi-Square computations generated above.  